Shared Subspace Model

### 2016-05-10 paper/mscov.tex
Updated paper.

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{multirow}
\usepackage[margin=.75in]{geometry}

@@

\bibliographystyle{plain}
\title{Shared Subspace Covariance Model}
\author{Alexander Franks and Peter Hoff}
\date{\today}
\maketitle

\begin{abstract}
  We develop model-based methods for evaluating similarities and
  differences among several $p\times p$ covariance matrices in the
  $p>n$ setting.  This is done by assuming a spiked covariance model
  for each group, and sharing information about the subspace spanned
  by the group-level eigenvectors.  We propose an EM algorithm for
  estimating the shared subspace and covariance matrix of the data
  projected onto this subspace.  In addition, we provide an Gibbs
  sampling algorithm for estimating the posterior uncertainty of
  group-level eigenvectors and eigenvalues on this subspace.

@@

%\citet{donoho_gavish_johnstone_2013} 
%\citet{paul_2007}

\section{The Spiked Covariance Model}  
Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$) 
distribution with density 
given by 
\begin{equation} 
p(\bl S | \Sigma, n) \propto_{\Sigma} l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
\label{eqn:lik}
\end{equation}
where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.

@@

Such a likelihood results from $\bl S$ being, for example, a residual 
sum of squares matrix from a multivariate regression analysis. In  this case,
$n$ is the number of independent observations minus the 
rank of the design matrix. 
The spiked principle 
components model (spiked PCA) introduced by \citet{Johnstone2001} assumes that 
%\begin{equation} 
$\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )$  ,  
%\end{equation}
where $\Lambda$ is an $r\times r$ diagonal matrix and 
$\bl U \in \mathcal V_{r,p}$, with 
%$r< ( n \wedge p)$. 
$r<p$. Such a formulation is appealing because it explicitly
partitions the covariance into a tractable low rank ``signal'' and
homscedastic ``noise''.

@@

However, in many modern applications, the primary objective is to
compare covariances across multiple groups of observations.  For
instance, Cite
Daniela Witten.  Hoff eigenpooling, no longer equivariance.  In suhc a
situation, equivariance blah blah.  In the following section, we
propose a model for estimating and exploring differences between
covariance matrices from multiple groups of related data.

\section{The Shared Subspace Spiked Covariance Model}

Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ which follow
the spiked covariance model with density as given in
\ref{eqn:rplik}.    Each group consists of of $n_k$ measurements of $p$ features, typically with $n_k << p$.  However, we assume that the eigenvectors
are related across groups in a particular way.  \citet{Hoff2009}
introduced a model for eigenvector shrinkage based on the matrix Bingham
distribution.

In contrast, in this paper, we assume a priori that the eigenvectors
are uniformally distributed on a low rank common subspace.  That is,

\begin{equation}
\Sigma_k = V\psi_kV^T + \sigma^2_kI
\end{equation}

With $V \in \mathcal{V}_{s,p}$ with $p
\geq s \geq r$.  Here, $VV^T \in \mathcal{G}_{s, p}$ is the plane of
variation shared by all groups and $\psi_k$ is the rank-r covariance matrix
on that subspace.  \textbf{Rationale}.

In this paper, as in blah we consider a model on eigendecomposition of
$\psi_k$ so that Equation \ref{} can be written as

\begin{equation}
\Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
\end{equation}

Where, $\Lambda_k$ is and $r \times r$ diagonal matrix of eigenvalues,
and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  Note that for any single group we can recover the original
spiked covariance formuation (Equation
\ref{}) by letting $U_k = VO_k$.

@@

\begin{equation}
p(S | \Sigma_k,n_k) \propto \prod_{k=1}^K |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
\end{equation}

@@

We can write the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ given $Y_k$ using Equation \ref{eqn:rplik} by replacing
$U_k$ with $VO_k$: 
\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto \prod_{k=1}^K\left[ (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki}) \right) ^{n_k/2}  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \right]
\end{equation}

\subsection*{An Expectation-Maximization Algorithm For Estimating the Shared Subspace}

We start by outlining a computationally efficient algorithm for
estimating the shared subspace, $VV^T$.

The full likelihood can be written as

\begin{align}
p(\bl S | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}

Where $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
$V$ (up to additive constant) is

To find the maximum marginal likelihood of $V$, we use an EM approach,
treating the $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing''
parameters to be imputed. %%We assume a Wishart prior on $M_k^{-1}$ and an Gamma prior
on $\frac{1}{\sigma_k^2}$.  If we take these prior's to be
``non-informative'' degenerate priors, we have

We put a Jeffreys prior on both $M_k^{-1}$ and $\frac{1}{\sigma_k^2}$.

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k + r + 1$).

\noindent For $\frac{1}{\sigma_k^2}$ we have

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

\noindent which is a Gamma($n_k(p-r)/2+1$, $\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$)

\noindent Thus, for each iteration, $t$, of the EM algorithm:

\begin{enumerate}
\item For each $k$, compute relevant conditional expectations:
\begin{enumerate}
\item $E[M_k^{-1} | V_{t-1}] = (n_k + r  + 1)(V^T \mathbf{S}_kV)^{-1}$
\item $E[\frac{1}{2\sigma_k^2}|V_{t-1}] = \frac{n_k(p-r) + 2}{tr((I-VV^T)S_k)}$
\end{enumerate}
\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{t-1}]V^T +
       E[\frac{1}{2\sigma_k^2}|V_{t-1}]VV^T\right)\mathbf{S}_k/2)$ (optimization
  approach from Wen et al, 2013)
\end{enumerate}

\subsection{Alternative formulation}
$M_k = \sigma_k^2(\Phi_k + I)$ where 
$$ \Phi_k =\left( \begin{array}{cc}
\Psi_k & 0  \\
0 & D  \end{array} \right)$$

Letting $V = [V_1\, V_2]$

\begin{eqnarray*}
p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
                & = (|\psi_k + I|^{-1}|D +
                  I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
                  -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
\end{eqnarray*}

$$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$

\section{Bayesian Inference}

\subsection{Prior and posterior distributions}

\paragraph{Conditional estimation of $(\sigma^2, \bl U)$:}
Bayesian inference is facilitated with the use of semiconjugate 
prior distributions  for $(\sigma^2, \bl U)$.
These include the inverse-gamma class of priors for $\sigma^2$, 
and the von Mises-Fisher-Bingham  
distributions for $\bl U$ \citep{hoff_2009a,hoff_2009b}.  
In the absence of real prior information,  invariance 
considerations suggest the use of priors on $\sigma^2$ and 
$\bl U$ that lead to equivariant estimators. 
In particular, an (improper) prior having density 
$p(\sigma^2,\bl U) = 1/\sigma^2 $, with respect to the 
product of Lebesgue measure on $\mathbb R^+$ and the 
uniform probability measure on $\mathcal V_{r,p}$, 
leads to Bayes estimators $\hat \Sigma(\bl S)$ that 
are equivariant with respect to 
rotations and scale changes, so that 
$\hat \Sigma(a \bl W \bl S \bl W^T) = 
   a \bl W \hat\Sigma(\bl S) \bl W$ for all 
$a>0$ and $\bl W\in \mathcal O_{p}$  
(assuming an invariant loss function).  
Under this invariant prior, straightforward calculations
show that the full conditional distribution of
$\sigma^2$ is
inverse-gamma$( n p/2 , \tr(\bl S(\bl I-\bl U\Omega\bl U)/2)$.
%Under the uniform prior on $\mathcal V_{r,p}$,
The full conditional distribution of $\bl U$ has a density proportional
to the likelihood, so
\[
 p(\bl U | \bl S, \sigma^2, \Omega) \propto 
  \etr( \Omega \bl U^T [\bl S/(2\sigma^2) \bl U). 
\]
This is a Bingham$(\Omega, \bl S/(2\sigma^2)$ distribution on
$\mathcal V_{r,p}$ \citep{khatri_mardia_1977}. A Gibbs sampler to
simulate from this distribution is given in
\citet{hoff_2009a}.
Interestingly, if $\Omega$ were known (which it is not), then 
for a given invariant loss function
the Bayes estimator under this prior 
minimizes the (frequentist) risk among all 
equivariant estimators \citep{eaton_1989}.

>>>>

\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow, booktabs}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{chemarrow}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{tikz}
\usepackage[margin=.75in]{geometry}

@@

\title{Shared Subspace Covariance Model}
\author{Alexander Franks and Peter Hoff}
\date{\today}
\maketitle

\begin{abstract}

% From Hoff Although the covariance matrices corresponding to different
% populations are unlikely to be exactly equal they can still exhibit a
% high degree of similarity. For example, some pairs of variables may be
% positively correlated across most groups, whereas the correlation
% between other pairs may be consistently negative. In such cases much
% of the similarity across covari- ance matrices can be described by
% similarities in their principal axes, which are the axes that are
% defined by the eigenvectors of the covariance matrices. Estimating the
% degree of across-popu- lation eigenvector heterogeneity can be helpful
% for a variety of estimation tasks.

  We develop model-based methods for evaluating similarities and
  differences among several $p\times p$ covariance matrices in the
  $p > n$ setting.  This is done by assuming a spiked covariance model
  for each group, and sharing information about the subspace spanned
  by the group-level eigenvectors.  We propose an EM algorithm for
  estimating the shared subspace and the covariance matrix of the data
  projected onto this subspace.  We also provide a Gibbs sampling
  algorithm for estimating the posterior uncertainty of group-level
  eigenvectors and eigenvalues.  Finally, we demonstrate the utility
  of our method in analyses exploring covariability in biological data.

@@

Introduction to covariance estimation and spiked covariance model.

Suppose $\bl S$ has a possibly degenerate Wishart$(\Sigma,n$) 
distribution with density 
given by 
\begin{equation} 
p(\bl S | \Sigma, n) \propto_{\Sigma} l(\Sigma:\bl S) =  |\Sigma|^{-n/2} \text{etr}( - \Sigma^{-1} \bl S/2 ) ,  
\label{eqn:lik}
\end{equation}
where $\Sigma \in \mathcal S_p^+$ and $n$ may be less than $p$.

@@

Such a likelihood results from $\bl S$ being, for example, a residual 
sum of squares matrix from a multivariate regression analysis. In  this case,
$n$ is the number of independent observations minus the 
rank of the design matrix. 
The spiked principle 
components model (spiked PCA) introduced by \citet{Johnstone2001} assumes that 
\begin{equation} 
\Sigma = \sigma^2 (  \bl U  \Lambda  \bl U^T  + \bl I )
\label{eqn:spiked}
\end{equation}

where $\Lambda$ is an $r\times r$ diagonal matrix and 
$\bl U \in \mathcal V_{r,p}$, with 
%$r< ( n \wedge p)$. 
$r<p$. Thus, the spiked covariance model is a low rank factor model
with homoscedastic error \citep{}.  Such a formulation is appealing because it explicitly
partitions the covariance into a tractable low rank ``signal'' and homscedastic ``noise''.

@@

However, in many modern applications, it is not enough to
independently estimate an individual covariance matrix.  Rather, a
comparative analysis of covariance patterns across multiple groups of
observations would be most informative.  For example, in case
control-studies, mean level effects may be small relative to subject
variability.  Differences in the covariance
matrices may lead to better insights about the differences between groups.
Alternatively, when mean level differences are detectable, better
estimates of the covariability of features across groups may lead to improved
understanding of the mechanisms underlying these apparent mean level
differences.

With this situation in mind, Flury (1984) developed estimation and
testing pro-cedures for the ‘common principal components’ model, in
which a set of covariance matrices
\citep{Flury1987} and \citep{Schott1991, Schott1999} considered
cases in which only certain columns or subspaces of U are shared across populations, and \citep{Boik2002} described a very general model in which eigenspaces can be shared between all or some of the populations.

Continuously parameterized covariance matrices. \citep{Hoff2011}, \citep{Chiu1996}, \citep{Yin2010}

In such models, it is no longer reasonable to assume orthogonal equivariance .
To address these issues, \citet{Hoff2009} introduced a model for
eigenvector shrinkage based on the matrix Bingham distribution.

Partial least squares discriminant analysis refs?
\citep{Westerhuis2008, Szymanska2012, Eriksson2001, Barker2003, Saebo2008}

Recently there has been significant interest gaussian graphical
models, .   In gaussian graphical models, zeros in the
precision matrix correspond to conditional independences between pairs
of features given the remaining features \citep{}.  \citet{Witten2014} developed a
method for pooled estimation inverse covariance matrices across
multiple groups.  Identifying a shared pattern of zero's in the inverse covariance
matrices for multiple classes of observations.

In the following section, we
propose a model for estimating and exploring differences between
covariance matrices from multiple groups of related data.

Below we describe our approach for comparaing covariance matrices on
a common low dimensional subspace.

\section{The Shared Subspace Spiked Covariance Model}
\label{sec:shared}

Assume we have $k=1, ..., K$ groups of matrices $\bl S_k$ each of
which follow a (degenerate) Wishart distribution, as in Equation
\ref{eqn:lik}.  Each group consists of of $n_k$ measurements of $p$
features, typically with $n_k << p$.  We pool information about the
covariance matrices by assuming that most of the relevant variation
for each group shares a low rank subspace spanned by
the columns of a matrix $V$.  That is,

\begin{equation}
\Sigma_k = V\psi_kV^T + \sigma^2_kI
\label{eqn:sspsi}
\end{equation}

Here, $V \in \mathcal{V}_{s,p}$ and $p \geq s$.  $V$ is only
identifiable up to right rotations.  Consequently,
$VV^T \in \mathcal{G}_{s, p}$, which defines the plane of variation
shared by all groups, is identifiable.  $\psi_k$ is a rank-$r$
covariance matrix on the $s$-dimensional subspace.  Such a model is
reasonable for many applications involving groups of highly structure
data on the same set of $p$ features.  For instance, in biological
applications, only certain subsets of molecules can physically
interact which constrains the space of possible correlation patterns.
Our model exploits this notion, by estimating differences between
groups on a common lower dimensional subpace.

\section{Inference}
\subsection{An Expectation-Maximization Algorithm}
\label{sec:em}

We start by outlining a computationally efficient algorithm for
estimating the shared subspace, $VV^T$, noise $\sigma_k^2$ and
projected covariance matrix $\psi_k$.

First, note that the full likelihood can be written as

\begin{align}
p(S_1, ... S_k | \Sigma_k,n_k) &\propto \prod_{k=1}^K |\Sigma_k|^{-n_k/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)  \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-(\sigma_k^2(V\psi_kV^T +
  I))^{-1}\mathbf{S}_k/2) \\
&\propto \prod_{k=1}^K  |\Sigma_k|^{-n_k/2}etr(-\left[V(\psi_k +
  I)^{-1}/\sigma_k^2 V^T + (I-VV^T)/\sigma^2_k\right]\mathbf{S}_k/2)
  \\
&\propto \prod_{k=1}^K  (\sigma_k^2)^{-n_k(p-r)/2}|M_k|^{-n_k/2}etr(-\left[VM_k^{-1}V^T + \frac{1}{\sigma^2_k} (I-VV^T)\right]\mathbf{S}_k/2) 
\end{align}

Where we define $M_k^{-1} = (\psi_k + I) ^{-1}/\sigma_k^2$.  The log-likelihood in
$V$ (up to additive constant) is

We propose a computationally efficient expectation-maximizatino (EM)
algorithm, to estimate the relevant quantities.  In particular we
maximize the marginal posterior distribution of $V$, treating the
$M_k^{-1}$ and $\frac{1}{\sigma_k^2}$ as the ``missing'' parameters to
be imputed.  We assume independent Jeffreys' prior distributions for
both $\frac{1}{\sigma_k^2}$ and $M_k^{-1}$ and a uniform prior for $V
\in \mathcal{V}_{s, p}$.  In particular we let
$p(\frac{1}{\sigma_k^2}) \propto \sigma_k^2$ and
$p(M_k^{-1}) \propto |M_k^{-1}|^{-(r+1)/2}$.  From the likelihood it
can easily be shown that the conditional posterior for $M_k^{-1}$ is

$$p(M_k^{-1} | V) \propto |M_k^{-1}|^{(n_k - r -1)/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) $$

\noindent which is a Wishart($V^T\mathbf{S}_kV$, $n_k$).  The
conditional posterior distribution of $\frac{1}{\sigma_k^2}$ is simply

$$p\left(\frac{1}{\sigma^2} | V\right) \propto \left(\frac{1}{\sigma_k^2}\right)^{n_k(p-r)/2-1}etr\left(-\frac{1}{2\sigma^2_k} (I-VV^T)\mathbf{S}_k\right)  $$

\noindent which is a Gamma($n_k(p-r)/2$,
$\frac{1}{2}tr((I-VV^T)\mathbf{S}_k)$).  Thus, for each iteration,
$t$, of the EM algorithm:

\begin{enumerate}
\item For each $k$, compute relevant conditional expectations:
\begin{enumerate}
\item $E[M_k^{-1} | V_{(t-1)}] = n_k(V_{(t-1)}^T \mathbf{S}_kV_{(t-1)})^{-1}$
\item $E[\frac{1}{2\sigma_k^2}|V_{(t-1)}] = \frac{n_k(p-r)}{tr((I-V_{(t-1)}V_{(t-1)}^T)S_k)}$
\end{enumerate}
\item Compute $V_{t}$ = $\underset{V}{argmax}  \sum_k tr\left(-VE[M^{-1}|V_{(t-1)}]^T +
       E[\frac{1}{2\sigma_k^2}|V_{(t-1)}]VV^T\right)\mathbf{S}_k/2)$ 
\end{enumerate}

For the ``M-step'', we use a numerical optimization algorithm based on
the Cayley transform to preserve the orthogonality constraints in $V$
\citep{Wen2013}.  Importantly, the complexity of this algorithm is
dominated by dimension of the shared subspace, $s$, not the number of
features $p$.  Thus, our approach is computationally efficient for
relatively small values of $s$, even when $p$ is large.

% \subsection{Alternative formulation}
% $M_k = \sigma_k^2(\Phi_k + I)$ where 
% $$ \Phi_k =\left( \begin{array}{cc}
% \Psi_k & 0  \\
% 0 & D  \end{array} \right)$$

% Letting $V = [V_1\, V_2]$

% \begin{eqnarray*}
% p(M_k^{-1} | V) &\propto \prod_k |M_k^{-1}|^{n_k/2}etr(-1/2(M_k^{-1}V^T\mathbf{S}_kV)) \\
%                 & = (|\psi_k + I|^{-1}|D +
%                   I|^{-1})^{n_k/2}etr(-1/2((\psi_k+I)^{-1}V_1^T\mathbf{S}_kV_1
%                   -1/2(D+I)^{-1}V_2^T\mathbf{S}_kV_2)) \\
% \end{eqnarray*}

% $$p(\frac{1/d_i} | V) \propto \prod_k (d_i + 1)^{-n_k/2}exp\left(\frac{1}{-1/2(d_i+1)}(V_2^T\mathbf{S}_kV_2)_{i,i}\right) $$

\subsection{Bayesian Inference}
\label{sec:bayes}

Although the EM algorithm presented in the previous section yields
point estimates of $\psi_k$, it does not lead to natural uncertainty
bounds for these estimates.  To quantify the uncertainty, we consider Bayesian
inference for $\psi_k$, given the
subspace $VV^T$.  Note that when the subspace is known, the posterior
distribution of $\Sigma_k$ is conditionally independent from the other
groups, i.e.  $P(\Sigma_{k} | V, \bl S_k, \Sigma_{-k}) = P(\Sigma_{k} | V, \bl S_k)$.
Thus, we can independently estimate the conditional posterior
distribution for each group.  In this paper, we focus on models based
on the eigendecomposition of $\psi_k$.  In this setting, each covariance
$\Sigma_k$ is a spiked covariance matrix where Equation \ref{eqn:sspsi}
can be written as

\begin{equation}
\Sigma_k = VO_k\Lambda_KO_k^TV^T + \sigma^2_kI
\label{eqn:ss}
\end{equation}

Here, $\Lambda_k$ is an $r \times r$ diagonal matrix of eigenvalues,
and $O_k \in \mathcal{V}_{s,r}$ is the matrix of eigenvectors of
$\psi_k$.  For any individual group, this corresponds to the original
spiked covariance model (Equation \ref{eqn:spiked}) with $U_k = VO_k$.

@@

\begin{equation}
p(S | \Sigma_k,n_k) \propto |\Sigma_k|^{-n/2}etr(-\Sigma_k^{-1}\mathbf{S}_k/2)
\end{equation}

@@

We can write the likelihood of $V$, $O_k$, $\Lambda_k$ and
$\sigma_k^2$ given $Y_k$ using Equation \ref{eqn:rplik} by replacing
$U_k$ with $VO_k$: 
\begin{equation}
 L(\sigma_k^2,\bl V , O_k \Omega_k : \bl Y_k) \propto
    (\sigma_k^2)^{-n_kp/2}etr(-\frac{1}{2\sigma_k^2}\mathbf{S}_k)\left(\prod_{i=1}^r(1-\omega_{ki})
   \right) ^{n_k/2}
   etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k)
\label{eqn:sslik}
\end{equation}

We use conjugate and semiconjugate priors for the parameters $O_k$,
$\sigma^2_k$ and $\Omega_k$ to facilitate inference.  In the absence
of real prior information, invariance considerations suggest the use
of priors that lead to equivariant estimators.  Below we describe our
choices for the prior distributions of each parameter and the
resultant posterior distributions.

\paragraph{Conditional distribution of $\sigma_k^2$:}

From Equation \ref{eqn:sslik} it is clear that the the inverse-gamma
class of prior distributions is conjugate for$\sigma_k^2$.  We chose a
default prior distriubtion for $\sigma^2$ that is equivariant with
respect to scale changes.  Specifically, we use Jeffreys' prior, an
improper prior with density $p(\sigma^2) \propto 1/\sigma^2 $.  Under
this prior, straightforward calculations show that the full
conditional distribution of $\sigma_k^2$ is
inverse-gamma$( n_k p/2 , \tr(\bl S_k(\bl I -\bl U_k\Omega_k\bl
U_k)/2)$, where $U_k = \hat{V}O_k$

\paragraph{Conditional distribution of $O_k$:} Given the likelihood
from Equation \ref{eqn:sslik}, it is easy to show that the class of
von Mises-Fisher-Bingham distributions are conjugate for $O_k$
\citep{Hoff2009, Hoff2012}.  Again, invariance considerations
lead us to use a rotationally invariant uniform probability measure on
$\mathcal V_{s,p}$.  Under this uniform prior, the full conditional
distribution of $\bl O_k$ has a density proportional to the
likelihood, so

\begin{align}
\label{lik_vo}
 p(O_k | \sigma^2_k, U_k, \Omega_k) & \propto etr(\Omega_kO^T_k\hat{V}^T[S_k/(2\sigma^2_k)]\hat{V}O_k)
\end{align}

\noindent This is a Bingham$(\Omega, \hat{V}^T \bl S_k \hat{V}/(2\sigma^2)$
distribution on $\mathcal V_{s, p}$ \citep{Khatri1977}. A
Gibbs sampler to simulate from this distribution is given in
\citet{Hoff2012}.

Together, the prior for $\sigma_k^2$ and $O_k$ leads to conditional
(on $V$) Bayes estimators $\hat \Sigma(V^T \bl S V)$ that are
equivariant with respect to scale changes and rotations on the
subspace spanned by $V$, so
that $\hat \Sigma(a \bl W V^T \bl S V \bl W^T) = a \bl W \hat\Sigma(V^T
\bl S V) \bl W$
for all $a>0$ and $\bl W\in \mathcal O_{s}$ (assuming an invariant
loss function). Interestingly, if $\Omega$ were known (which it is
not), then for a given invariant loss function the Bayes estimator
under this prior minimizes the (frequentist) risk among all
equivariant estimators \citep{Eaton1989}.

\paragraph{Conditional distribution for $\Omega$:} Here we specify the conditional
distribution of the diagonal matrix $\Omega_k =
\Lambda_k(I+\Lambda_k)^{-1} = \text{diag}(\omega_{k1}, ... \omega_{kr})$.  We first consider a uniform(0,1) prior distribution for each element of $\Omega$, or
equivalently, an $F_{2,2}$ prior distribution for the elements of
$\Lambda$.  The full conditional distribution of an
element $\omega_i$ of $\Omega$ is proportional to the likelihood
function, so

\begin{align}
p(\omega_{ki}|V, O_k, \bl S_k) \propto_{\omega_{ki}}
  \left(\prod_{i=1}^r(1-\omega_{ki})^{n_k/2}  \right)
  etr(\frac{1}{2\sigma_k^2}(VO_k\Omega_kO_k^TV^T)\mathbf{S}_k) \\
&  \propto & (1-\omega_{ki})^{n/2} e^{c \omega_s  n/2},    
\label{eqn:wpost}
\end{align}

where $c = \bl u_{ki}^T S_k \bl u_{ki}/(n \sigma^2)$ and $\bl u_{ki}$ is
column $i$ of $U_k = VO_k$.  While not proportional to a density
belonging to a standard class of distributions, we can sample from the
corresponding univariate distribution numerically.  The behavior of
this distribution is straightforward to understand: if $c\leq 1$, then
the the function has a maximum at $\omega_i =0$, and decays
monotonically to zero as $\omega \rightarrow 1$.  If $c>1$ then the
function is uniquely maximized at $(c-1)/c \in (0,1)$.  To see why
this makes sense, note that the likelihood is maximized when the
columns of $\bl U_k$ are equal to the eigenvectors of $\bl S_k$
corresponding to its top $r$ eigenvalues
\citep{Tipping1999}. At this value of $\bl U_k$, $c$ will then
equal one of the top $r$ eigenvalues of $\bl S_k/(n\sigma_k^2)$.  In the
case that $n\gg p$, we expect
$\bl S_k/(n_k\sigma_k^2)\approx \Sigma_k/\sigma_k^2$, the true (scaled)
population covariance, and so we expect $c$ to be near one of the top
$r$ eigenvalues of $\Sigma_0/\sigma^2$, say $\lambda_0+1$.  If indeed
$\Sigma_k$ has $r$ spikes, then $\lambda_k>0$,
$c \approx \lambda_0 +1 > 1$, and so the conditional mode of $w_i$ is
approximately $(c-1)/c = \lambda_i/(\lambda_i+1)$, the correct value.
On the other hand, if we have assumed the existence of a spike when
there is none, then $\lambda_0=0$, $c\approx 1$ and the Bayes estimate
of $w_i$ will be shrunk towards zero, as it should be.

% \paragraph{Conditional distribution of $V$}
% Need to finish / should we include?

% Although we prefer EM... it is also posible to estimate $V$ using
% Bayesian inference.  We assume a uniform prior on $V \in \mathcal{V}_{p, s}$ so that the
% conditional posterior of $V$ is proportional to the likelihood:

% \begin{align}
% p(V|\bl O, \bl S, \bl \sigma^2, \bl \Omega) &\propto etr(\sum_{k=1}^K[S_k/(2\sigma^2_k)]V[O_k\Omega_kO^T_k]V^T)\\
% \end{align}

% Although close in appearance, this conditional distribution is not a matrix
% Bingham-von Mises-Fisher distribution.  However, the conditional distribution
% of a column of $V$ given all other columns is a vector Bingham-von
% Mises-Fisher distribution.  To see this, let
% $A_k = S_k/(2\sigma^2_k)$ and $B_k = O_k\Omega_kO^T_k$.  We can write
% the conditional distribution of $V$ given $A_k$ and $B_k$ as
% \begin{align}
% P(V|A_k,B_k) &\propto etr(\sum_{k=1}^KA_kVB_kV^T)\\
% &= etr(\sum_{k=1}^K\sum_{i,j}^sA_kv_iv_j^Tb_{ijk})\\
% &= etr(\sum_{i,j}^s \sum_{k=1}^K v_j^T[b_{ijk}A_k]v_i)
% \end{align}

% From here, we specify the conditional distribution of a single column of $V$ given $A_k$, $B_k$ and all other columns of V:

% \begin{align}
% P(v_i|A_k,B_kV_{-i})  &\propto etr(2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])v_i+v_i^T[\sum_{k=1}^K b_{ijk}A_k]v_i))\\
% &= etr(Cv_i+v_i^TMv_i)
% \label{eqn:vbmf}
% \end{align}

% This is a vector Bingham-von Mises-Fisher Distribution, BMF(C, M)
% with $C = 2(\sum_{j\neq i}^s v_j^T[\sum_{k=1}^K b_{ijk}A_k])$ and $M=
% \sum_{k=1}^K b_{ijk}A_k$.  Following \citet{Hoff2012},  we can derive
% a Gibbs sampler for $V$ based on conditional draws of the columns of
% $V$.
